\def\filepath{templates}
%\def\filepath{C:/Users/holden-lee/Dropbox/Math/templates}

\input{\filepath/packages_book.tex}
\input{\filepath/theorems_with_boxes.tex}
\input{\filepath/macros.tex}
\input{\filepath/formatting.tex}
\input{\filepath/other.tex}
\input{\filepath/theorem_num.tex}

%\newcommand{\Bin}{\operatorname{Bin}}
\def\name{Stochastic processes on graphs}


\pagestyle{fancy}
%\addtolength{\headwidth}{\marginparsep} %these change header-rule width
%\addtolength{\headwidth}{\marginparwidth}
\lhead{MAT589}
\chead{} 
\rhead{Stochastic processes on graphs} 
\lfoot{} 
\cfoot{\thepage} 
\rfoot{} % !! Remember to change the problem set number
\renewcommand{\headrulewidth}{.3pt} 
%\renewcommand{\footrulewidth}{.3pt}
\setlength\voffset{0in}
%\setlength\textheight{648pt}

\begin{document}
%\input{\filepath/titlepage.tex}
%\maketitle
\title{Stochastic processes on graphs}
\author{Lectures by Professor Allan Sly\\ Scribe: Holden Lee}
\maketitle


\tableofcontents

%\startcontents
%\printcontents{ }{-1}{}

\chapter*{Introduction}

%Princeton, Spring 2018. 

We will cover:
\begin{enumerate}
\item
Random graphs
\begin{enumerate}
\item
Classical results
\item
Erd\H os-Renyi random graph
\item
Properties (connectivity, components, local structures)
\end{enumerate}
\item
Stochastic processes/interacting particle systems
\begin{enumerate}
\item
Spin systems (Ising or hard-core model)value 
\item
Infection, contact process, SIS, SIR
\item
Diffusion of information, cascades
\end{enumerate}
\item
Random constraint satisfaction problems (CSPs). The canonical example is:
\begin{enumerate}
\item
Coloring a random graph with $k$ colours.

Here the variables are colours of vertices and the constraints are the edges: neighbours have different colours.

The more constraints, the harder. 
\end{enumerate}
Statistical mechanics give a lot of predictions but not proofs.

As density increases, there are fewer colorings. At first there is one large component; you can get from any coloring to any other through a path of valid colorings. 

After a threshold, the component splits up into many clusters. This transition also seems to be the point where all the efficient algorithms (polytime in number of vertices) for finding a coloring break down. %The best ones all break here.

Then there's another phase transition after which you have no more colorings.

There is also a middle transition that tells you how many colorings you have.
Not many of these predictions have actually been proved; there are just bounds.

There are similar transitions for other problems like random $k$-SAT, largest independent set.
\end{enumerate}

%\setcounter{chapter}{1}
\input{2-5.tex}
\input{2-7.tex}
\input{2-14.tex}
\input{2-19.tex}
\input{2-21.tex}
\input{2-26.tex}
\input{2-28.tex}
\input{3-5.tex}
%INSERT_HERE


%\appendix
%\input{distribution_chapters/a.tex}

%\bibliographystyle{plain}
%\bibliography{\filepath/refs}

\printnomenclature
\printindex
\end{document}